### 2022-09-26 Thesis/text/einleitung.tex
grammarly

Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)

Description of the Problem

Robots regurlarly needs operator in case smth unexpected happens\\

this may include: unexpected collisions, a system restart is needed, battery is empty\\

The system is not able to react and handle unforeseen situations\\

This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.\\

Due to the need of the robot to require an operator, the robot is closer to an automated instead of an autonomous vehicle. 
Difference is defined by the SAE Levels for automonous driving in cars.\\

It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.\\

In theory the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem. And even when failures inside this subsystems are discovered, the options to handle these problems are very limited and often do not deal with the problems in an autonomous way. The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals.

Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments. Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.

In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot.

This thesis proposes a comprehensive approach for behavior planning for an autonomous mobile robot.

\textbf{How can behavior planning increase the safety and autonomy of an autonomous mobile robot running ROS?}

>>>>

Mobile robots are becoming more widespread in many domains. They are marketed as autonomous mobile robots, but their autonomy is often limited to particular environments and requires frequent human help to function correctly. This reliance on a human operator hardly qualifies the robot to be labeled autonomous. 
\subparagraph*{}
One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS). ROS offers many tools and functionalities to create and program robots, but a standard robot needs human supervision during the operation to ensure proper function and safety. 
There are ROS packages to safely navigate from point A to point B, but currently, typical robots cannot react to unforeseen events. The robot needs to be able to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment. These can be either other robots or humans. 
\subparagraph*{}
In theory, the robot can perform fully autonomous navigation, including mapping and localization. However, when something unexpected happens, the robot's autonomy can not be guaranteed any more. 
A default robot cannot represent all possible fail states and does not detect failures on a systematic level but only inside its navigation-related subsystems. Moreover, when failures inside these subsystems are discovered, the options to handle these problems are minimal and often do not deal with the problems in an autonomous way. The reliance on external problem-solving decreases the robot's usefulness when tasked with tangible goals. 
Therefore, this thesis centers around the research question of how behavior planning can increase the robustness and autonomy of a robot running ROS2. 
Behavior planning enables the robot to react to failures and problems of the system and can decide alternative courses of action to mitigate risks and finish the given tasks.

\subparagraph*{}
This thesis will explore possibilities for improving current systems by adding a dedicated behavior planning component. Furthermore, the thesis will provide an implementation of exemplary behaviors and provide a system architecture to incorporate behavior planning in other robots, too.

\subparagraph*{}
The desired outcomes of this implemented software will be an increase in robustness and a decrease in required human actions during several scenarios. In these scenarios, failures and problems are artificially induced to test the robot's abilities to behave autonomously.

% BEFORE GRAMMARLY BELOW

%\chapter{Introduction}
%\label{cha:introduction}
%
%Mobile robots are becoming more widespread in many domains. They are marketed as autonomous mobile robots, but often the autonomy of the robot is limited to very specific environments and require frequent human help to function correctly. This hardly qualifies the robot to be labeled autonomous. 
%\subparagraph*{}
%One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS). ROS offers many tools and functionalities to create and program robots, but even then a normal robot needs human supervision during the operation to ensure the proper function and safety. 
%There are ROS packages to safely navigate from point A to point B, but currently they are lacking the ability to react to unforeseen events. It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans. 
%\subparagraph*{}
%In theory, the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore. 
%Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem. And even when failures inside this subsystems are discovered, the options to handle these problems are very limited and often do not deal with the problems in an autonomous way. The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals. 
%Therefore, this thesis centers around the research question how behavior planning can increase the robustness and autonomy of a robot running ROS2. 
%Behavior planning enables the robot to react to failures and problems of the system and can decide alternative courses of action to mitigate risks and finish the given tasks.
%
%%Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
%%
%%Description of the Problem:
%%
%%Despite the advancements in autonomous driving the robot regurlarly needs an operator in case smth unexpected happens.
%%
%%This may include: unexpected collisions, a system restart is needed, battery is empty, the environment is too crowded and the robot gets stuck
%%
%%The system is not able to react and handle unforeseen situations.
%%
%%This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.
%%
%%Due to the need of the robot to require an operator for safe, the robot is not really autonomous despite it being labeled as such. 
%%
%%(Difference is defined by the SAE Levels for automonous driving in cars.) \\
%%
%%Relevance, need for research:\\
%%\\
%
%%
%%Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments. Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.
%%
%%In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot. 
%%
%%Planned solution:\\
%\subparagraph*{}
%This thesis will explore possibilities how current systems can be improved through the addition of a dedicated behavior planning component. Furthermore, the thesis will provide an implementation of examplatory behaviors and provide a system architecture to incorporate behavior planning in other robots, too
%
%%The thesis will explore how an extension of the current software architecture and behavior planning capabilities for an autonomous mobile robot that is running ROS2 can improve the autonomy of the robot. 
%%
%%Expected results:\\
%\subparagraph*{}
%The desired outcomes of these implemented software will be an increase of robustness and decrease of required human actions during a number of scenarios. In these scenarios, failures and problems are artificially induced to test the robots abilities to behave autonomously. 
%
%
%%\textit{Research gap is not named so far}

### 2022-09-26 Thesis/text/ergebnisse.tex
grammarly

@@

>>>>

The robot is put into specifically designed scenarios that should trigger behavior responses from the behavior tree component. The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in chapter 3.2.

The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}. 
\begin{center}
\begin{figure}[ht]
	\includegraphics[width=0.75\textwidth]{images/house_env.png}
	\caption{Apartment Environment in Gazebo}
	\label{fig:house_gazebo}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
	\includegraphics[width=0.75\textwidth]{images/world_env.png}
	\caption{Enclosed Environment in Gazebo}
	\label{fig:world_gazebo}
\end{figure}
\end{center}

The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.

The robot location and navigation goals are defined in another ROS Node, which programmatically spawns the robot into the environment and sets the goal this way. This way, the same scenario can be run with and without behavior planning. To test the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in chapter 4.5.1, was shut down to emulate the sensor's failure.

@@

The scenarios are derived from the functional requirements and can be used to test multiple requirements as listed in the tables \ref{tab:sensor_scenarios} and \ref{tab:behavior_scenarios}.
The acceptance criteria often contain multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met for the test to be counted as a success.

For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario. 
Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered. The BT has to recover from the inability to plan in this scenario.
NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal. This scenario tests the ability to find alternative goals and navigate to a nearby goal. 
In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered. The forced collision tests the ability to escape a collision and resume navigation afterward. A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
Collision$\_$2 scenario is testing the same behavior but during active navigation. A flat obstacle is spawned in the path of the robot. The obstacle is too low to be detected by the laserscanner, and a collision occurs. The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario. 
The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge. The simulated battery is emptied to a low level, and a new navigation goal is set. If the battery charge runs below zero percent during the navigation, the test fails.

\begin{table}[ht]
	\caption{Sensor Scenarios}
	\label{tab:sensor_scenarios}
	\begin{tabular}{| m{0.1\textwidth} | m{0.17\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
  	\hline
  	\textbf{Name} & \textbf{Related Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
  	\hline
  	Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
  	\hline
  	Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
  	IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
  	\hline
  	IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
  	Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
  	\hline
  	Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
 	\end{tabular}
\end{table}

\begin{table}[ht]
	\caption{Behavior Scenarios}
	\label{tab:behavior_scenarios}
	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
	\hline
	\textbf{Name} & \textbf{Related Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
  	\hline	
  	NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still &  
Reset system, calculate path to goal, reach goal \\
	\hline
	NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
	\hline
	Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
	\hline
	Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) &  
Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
	\hline
	Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
	\hline
	Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
	\hline

	\end{tabular}
\end{table}

\section{Results}

The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for most test cases. The system supervision component is very reliable and enables the system to recover from an unexpected sensor failure. The more advanced behaviors implemented can improve the autonomous handling of the scenarios. The Battery Scenario was successful every time it was induced. However, the path planning behavior for finding alternative goals was unsuccessful about 50 percent of the time. The test showed that alternative methods for finding new goals would probably have resulted in better performance. The collision behavior during driving worked well, and the map updates are an essential feature for the success of the scenarios.
Nevertheless, there is a lack of intelligence and autonomy when the robot stands still. A collision is forced, but the implemented behavior relies on the same mechanism as during the driving scenario. The missing differentiation led to a test case where the robot reversed into a wall because it was turned 90 degrees by the collision. That test failed because a human operator would be needed to correct the robot. This lack of robustness was not apparent in the driving collision scenario because the robot's force was insufficient to cause a significant shift in orientation when colliding with obstacles.

\begin{table}[ht]

	\caption{Results}
	\label{tab:results}
	\begin{tabular}{| m{0.11\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.33\textwidth}|} 
  	\hline
  	\textbf{Name} & \textbf{Number of Runs} &  \textbf{Percentage Succesful Without Behavior Planning} & \textbf{Percentage Succesful With Behavior Planning}\\ 
  	\hline
  	Lidar$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	Lidar$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	Lidar$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	NoPath Found$\_$1 & 20 & 0 & 100 \\ 
  	\hline
  	NoPath Found$\_$2 & 20 & 0 & 55 \\ 
  	\hline
  	Collision$\_$1 & 20 & 0 & 95 \\ 
  	\hline
  	Collision$\_$2 & 20 & 0 & 100 \\ 
  	\hline
  	Battery$\_$1 & 10 & 0 & 100 \\ 
  	\hline
  	Motor$\_$1 & 10 & 0 & 0 \\ 
  	\hline
	\end{tabular}
\end{table}

The architectural design choices meet all the non-functional requirements. 
Implementing the whole behavior planning and system supervision system eliminates a single point of failure for the system. Even when the Navigation2 system collapses, is the robot able to move to a degree and, more importantly, stop completely (compare table \ref{tab:tab:nonfn_req}, non$\_$fn$\_$req). The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3). The robots' behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors use a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4). 
Finally, the system's performance when checking the conditions of the BT is below the desired threshold of 10ms, as the whole system runs in multiple threads, which allows the behavior tree to complete one complete execution cycle in about 5ms with the current system. (non$\_$fn$\_$req2).

### 2022-09-26 Thesis/text/methoden.tex
grammarly

\begin{table}[ht]
	\label{tab:ros_interfaces}
	\caption{ROS Communication Options}
	\begin{tabular}{ | m{0.15\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
  	\hline
  	Name & Communication Pattern & Area of Use & Cardinality \\ 
  	\hline
  	Topics & Publisher/ Subscriber & Continuous stream of data e.g. sensor data &  n:m \\ 
  	\hline
  	Services & Request/ Response & Get specific data only once & One service, many clients \\ 
  	\hline
  	Actions & Request/ Response and Publish/ Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
  	\hline
	\end{tabular}
\end{table}

@@

\begin{figure}[ht]
	\label{fig:nav_architecture}
	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
	\caption{Navigation2 Architecture \cite{macenski2020}}
\end{figure}
\subparagraph*{}
These packages are often integrating a SLAM method to build or enlarge maps. Path planning and execution capabilities, which correspond to the global and local planner described in section 2.2, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution. The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.

\begin{figure}[ht]
	\label{fig:bt_nav}
	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
	\caption{Navigation2 Behavior Tree}
\end{figure}
\subparagraph*{}
The behavior tree depicted in figure \ref{fig:bt_nav} is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map). These behaviors are executed if the robot can not make progress towards the set goal. 
Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way. The lifecycle approach allows the system degree to monitor the system node's startups, executions, and failures. 
\subparagraph*{}
With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy. The software can be modified and expanded with plugins to fit the users' needs. 
When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure \ref{fig:autonomous_driving_architectures}, the things that Nav2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in chapter 2.3.
\subparagraph*{}
The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably. One of the core problems behavior-based robots face is generating a dependable environment representation. An accurate representation is the basis for the last steps in the planning part of the "Sense - Think - Act" cycle. Unforeseen events can severely limit the reliability of the environment representation without the robot being aware of it. Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles. 
%\textit{Add statistics about random sensor failures every ~10000h} 
This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement. By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors. However, the deliberative behavior's planning is based upon a false representation of reality, leading to unsafe commands. The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior. Even worse would be that the reactive behaviors may not get activated, and the robot continues to drive despite being unaware of the surroundings. The unsafe behavior requires a human operator would need to step in and restore the robot to full functionality by hand.

@@

\begin{figure}[ht]
    \label{fig:turtlebot}
	\includegraphics[width=0.7\textwidth]{images/turtlebot_sim.png}
	\caption{Turtlebot3 model in Gazebo simulator}
\end{figure}

@@

The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous chapter.

\begin{table}[ht]
\label{tab:fn_req}
\caption{Functional Requirements}
	\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.62\textwidth}|} 
  	\hline
  	Nr. & Name & Priority & Description/Acceptance Criteria \\ 
  	\hline
  	fn$\_$req1 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information \\ 
  	\hline
  	fn$\_$req2 & Emergency Detection & High & The system detects emergency. Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
  	\hline
  	fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
  	\hline  	
  	fn$\_$req4 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
  	\hline
  	fn$\_$req5 & Maintain operability & High & The robot executes commands as long as it is safe. Ensure that the robot keeps driving if it is safe even when system functions are not working correctly. \\
  	\hline
  	fn$\_$req6 & Recovery & High &  The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
  	\hline  
  	fn$\_$req7 & Control Path Planning & Medium & The system controls and rates the quality of the planned paths. \\
  	\hline	
  	fn$\_$req8 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
  	\hline
  	fn$\_$req9 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal.  \\	
  	\hline
	\end{tabular}
\end{table}

\begin{table}[ht]
	\caption{Non-Functional Requirements}
	\label{tab:nonfn_req}
	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.08\textwidth} | m{0.56\textwidth} |} 
  	\hline
  	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description} \\ 
  	\hline
  	non$\_$freq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
  	\hline
  	non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
  	\hline
  	non$\_$freq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
  	\hline
  	non$\_$freq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
  	\hline
	\end{tabular}
\end{table}

@@

\begin{table}[ht]
	\caption{Control Nodes in BT.CPP}
	\label{tab:control_nodes_bt}
	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
  	\hline
  	Name & Child returns Failure & Child Returns Running \\ 
  	\hline
  	Sequence & Restart & Tick Again\\ 
  	\hline
  	Reactive Sequence & Restart & Restart \\ 
  	\hline
  	Sequence Star & Tick again & Tick again \\
  	\hline
  	Fallback & Tick next & Tick again\\
  	\hline  	
  	Reactive Fallback & Tick next & Restart \\
  	\hline
	\end{tabular}
\end{table}

@@

\begin{center}
\begin{table}[ht]
	\caption{Decorator Nodes in BT.CPP}
	\label{tab:decorators_bt}
	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
  	\hline
  	Name & Succeeds & Fails & Running \\ 
  	\hline
  	Inverter Node & If the child returns false & If child return success &  If the child returns running \\ 
  	\hline
  	ForceSuccess Node & Always & Never & If child returns running \\ 
  	\hline
  	ForceFailure Node & Never & Always & If child returns running\\
  	\hline
  	Repeat Node & Ticks child as long as it returns success. The maximum number of repeats can be defined & If the child returns failure & If the child returns running \\
  	\hline  	
  	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running. \\
  	\hline
	\end{tabular}
\end{table}
\end{center}

>>>>

\begin{table}[ht]
	\caption{ROS Communication Options}
	\label{tab:ros_interfaces}
	\begin{tabular}{ | m{0.15\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
  	\hline
  	\textbf{Name} & \textbf{Communication Pattern} & \textbf{Area of Use} & \textbf{Cardinality} \\ 
  	\hline
  	Topics & Publisher/ Subscriber & Continuous stream of data e.g. sensor data &  n:m \\ 
  	\hline
  	Services & Request/ Response & Get specific data only once & One service, many clients \\ 
  	\hline
  	Actions & Request/ Response and Publish/ Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
  	\hline
	\end{tabular}
\end{table}

@@

\begin{figure}[ht]
	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
	\caption{Navigation2 Architecture \cite{macenski2020}}
	\label{fig:nav_architecture}
\end{figure}
\subparagraph*{}
These packages are often integrating a SLAM method to build or enlarge maps. Path planning and execution capabilities, which correspond to the global and local planner described in section 2.2, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution. The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.

\begin{figure}[ht]
	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
	\caption{Navigation2 Behavior Tree}
	\label{fig:bt_nav}
\end{figure}
\subparagraph*{}
The behavior tree depicted in figure \ref{fig:bt_nav} is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map). These behaviors are executed if the robot can not make progress towards the set goal. 
Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way. The lifecycle approach allows the system degree to monitor the system node's startups, executions, and failures. 
\subparagraph*{}
With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy. The software can be modified and expanded with plugins to fit the users' needs. 
When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure \ref{fig:autonomous_driving_architectures}, the things that Nav2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in chapter 2.3.
\subparagraph*{}
The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably. One of the core problems behavior-based robots face is generating a dependable environment representation. An accurate representation is the basis for the last steps in the planning part of the "Sense - Think - Act" cycle. Unforeseen events can severely limit the reliability of the environment representation without the robot being aware of it. Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles. 
%\textit{Add statistics about random sensor failures every ~10000h} 
This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement. By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors. However, the deliberative behavior's planning is based upon a false representation of reality, leading to unsafe commands. The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior. Even worse would be that the reactive behaviors may not get activated, and the robot continues to drive despite being unaware of the surroundings. The unsafe behavior requires a human operator would need to step in and restore the robot to full functionality by hand.

@@

\begin{figure}[ht]   
	\includegraphics[width=0.7\textwidth]{images/turtlebot_sim.png}
	\caption{Turtlebot3 model in Gazebo simulator}
	\label{fig:turtlebot}
\end{figure}

@@

The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous chapter. 
The functional requirements and their acceptance criteria are listed in \ref{tab:fn_req}. The non-functional requirements are listed seperately in \ref{tab:nonfn_req}. 
To better contextualize the context of the systems requirements, figure \ref{fig:use_case} depicts a Use-Case diagram with the needed functions to improve the system's autonomy.

\begin{table}[ht]
\caption{Functional Requirements}
\label{tab:fn_req}
	\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
  	\hline
  	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description/Acceptance Criteria} \\ 
  	\hline
  	fn$\_$req1 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information \\ 
  	\hline
  	fn$\_$req2 & Emergency Detection & High & The system detects emergency. Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
  	\hline
  	fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
  	\hline  	
  	fn$\_$req4 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
  	\hline
  	fn$\_$req5 & Maintain operability & High & The robot executes commands as long as it is safe. Ensure that the robot keeps driving if it is safe even when system functions are not working correctly. \\
  	\hline
  	fn$\_$req6 & Recovery & High &  The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
  	\hline  
  	fn$\_$req7 & Control Path Planning & Medium & The system controls and rates the quality of the planned paths. \\
  	\hline	
  	fn$\_$req8 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
  	\hline
  	fn$\_$req9 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal.  \\	
  	\hline
	\end{tabular}
\end{table}

\begin{table}[ht]
	\caption{Non-Functional Requirements}
	\label{tab:nonfn_req}
	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.1\textwidth} | m{0.55\textwidth} |} 
  	\hline
  	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description} \\ 
  	\hline
  	non$\_$freq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
  	\hline
  	non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
  	\hline
  	non$\_$freq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
  	\hline
  	non$\_$freq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
  	\hline
	\end{tabular}
\end{table}

\begin{figure}[ht]   
	\includegraphics[width=0.7\textwidth]{images/use_case.png}
	\caption{Use Case Diagram}
	\label{fig:use_case}
\end{figure}

@@

\begin{table}[ht]
	\caption{Control Nodes in BT.CPP}
	\label{tab:control_nodes_bt}
	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
  	\hline
  	\textbf{Name }& \textbf{Child returns Failure} & \textbf{Child Returns Running} \\ 
  	\hline
  	Sequence & Restart & Tick Again\\ 
  	\hline
  	Reactive Sequence & Restart & Restart \\ 
  	\hline
  	Sequence Star & Tick again & Tick again \\
  	\hline
  	Fallback & Tick next & Tick again\\
  	\hline  	
  	Reactive Fallback & Tick next & Restart \\
  	\hline
	\end{tabular}
\end{table}

@@

\begin{center}
\begin{table}[ht]
	\caption{Decorator Nodes in BT.CPP}
	\label{tab:decorators_bt}
	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
  	\hline
  	\textbf{Name} & \textbf{Succeeds} & \textbf{Fails} & \textbf{Running} \\ 
  	\hline
  	Inverter Node & If the child returns false & If child return success &  If the child returns running \\ 
  	\hline
  	ForceSuccess Node & Always & Never & If child returns running \\ 
  	\hline
  	ForceFailure Node & Never & Always & If child returns running\\
  	\hline
  	Repeat Node & Ticks child as long as it returns success. The maximum number of repeats can be defined & If the child returns failure & If the child returns running \\
  	\hline  	
  	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running. \\
  	\hline
	\end{tabular}
\end{table}
\end{center}
